---
title: "Graphing"
author: Hadi Ahmad
thanks: "Code and data are available at: https://github.com/hadi-u/lecture-starter"
date: today
date-format: long
abstract: "This paper illustrates various types of graphs and themes using ``ggplot2``."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```


# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

We are interested in the height of Lake Huron (@fig-huronheight).

### Activity 2

```{r}
#| label: fig-huronheight
#| fig-cap: Annual measurements of the level, in feet, of Lake Huron 1875–1972.
#| echo: false

tibble(year = 1875:1972,
       level = as.numeric(datasets::LakeHuron)) |>
  ggplot(aes(x = year, y = level)) +
  geom_point() +
  labs(x = "Year",
       y = "Height (feet)") +
  theme_minimal()
```

### Activity 3
```{r}
#| label: fig-huronheight-bar
#| fig-cap: Annual measurements of the level, in feet, of Lake Huron 1875–1972.
#| echo: false

datasets::trees |> 
  as_tibble() |> 
  ggplot(aes(x = Height)) +
  geom_bar()


```


Some of our data is of penguins (@fig-bills), from @palmerpenguins.

### Activity 4

```{r}
#| label: fig-weight
#| fig-cap: Weight of chicks after birth
#| echo: false

datasets::ChickWeight |> 
  as_tibble() |> 
  ggplot(aes(x = Time, y = weight, group = Chick)) +
  geom_line() +
  theme_classic()

```


```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()

tibble(year = 1700:1988,
       sunspots = as.numeric(datasets::sunspot.year) |> round(0)) |>
  ggplot(aes(x = sunspots)) +
  geom_histogram() +
  theme_classic()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 

### Activity 6

```{r}
#| label: fig-planes2
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
#| 
ggplot(data = penguins,
       aes(x = flipper_length_mm,
           y = body_mass_g)) +
  geom_point(aes(color = species,
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkgray", "darkgray", "darkgray")) +
  labs(
    title = "According to all known laws of aviation, there is no way a penguin should be able to fly. Its wings are too small to get its fat little body off the ground. The penguin, of course, flies anyway because penguins don't care what humans think",
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Penguin species",
    shape = "Penguin species"
  ) +
  theme_dark() +
  theme(
    legend.position = c(0.5, 0.7),
    plot.title.position = "plot",
    plot.caption = element_text(hjust = 100, face = "bold"),
    plot.caption.position = "plot"
  )

```
```{r}

datasets::morley |> 
  tibble()


```


# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


